Improving Usability of Remote Access to Sensitive Data

### 2014-07-27 ccsw14/paper.tex
finished first draft of paper

\usepackage{epstopdf}
\usepackage{datetime}
\usdate
\settimeformat{ampmtime}

@@

\section{Implementation}

@@

%mode switch seamlessness (across machines)

%managing migration speeds for performance & network

@@

@@

%evaluation plans and some preliminary results

@@

Interdum et malesuada fames ac ante ipsum primis in faucibus. Nulla facilisi.
Etiam viverra condimentum accumsan. Mauris sed dolor sed mi dignissim
vulputate. Duis \cite{clark:pct} pharetra purus purus, eu ultricies ligula
dignissim vel. Sed nec mauris a turpis luctus placerat id ut ipsum. Fusce quis
tincidunt dolor.  Nunc mattis ornare sem, vel consectetur ipsum iaculis ac.
Aenean nulla lacus, iaculis sed tempor nec, euismod suscipit dui.

>>>>

\usepackage{epstopdf}
\usepackage{datetime}

% Switch between these two to add/remove todo comments
\usepackage{color}
\newcommand{\todo}[2]{{\color{red}\textbf{#2}}}
%\newcommand{\todo}[2]{#1}

\usdate
\settimeformat{ampmtime}

@@

\section{Implementation}
\label{sec:Implementation}

@@

As mentioned in Section \ref{sec:Implementation}, the migration-based mode
switch has currently been implemented as a demonstration, and has yet to be
fully integrated into the prototype.  We are currently addressing two
issues that will help usability and performance of this key element.

The first of these issues is the fact that migrating a VM across two physical
machines currently requires the user to reestablish a remote desktop or shell
connection with the destination machine.  We intend to set up a proxy connection
system that will redirect the user's connection automatically, without the need
for any action on the user end.  Such a proxy would need to reside on a machine
separate from machines serving VMs, otherwise it would create an intermachine
covert channel, since there would at some point be a connection link between a
process on a maintenance machine and a process on a secure machine.

The second issue with designing a mature migration-based mode switch is that of
speed and performance.  Transferring the state of many machines all running in
maintenance mode could easily strain even a very robust network.  Therefore it
is likely that some form of rate limiting will need to be imposed in order to
relieve network pressure.  The consequence of this is of course that the time it
takes to switch modes will increase.  We therefore plan to investigate to find
the best compromise between the two needs of limiting use of network bandwidth
and maximizing switch performance.

@@

Since the ultimate goal of the Data Capsules system is to improve usability of
remote access to sensitive data, it is natural that performance is an important
measure of its success.  Virtual machines must not suffer any slowdown, and the
new mode switching interface that it introduces must not be inconveniently slow.
Additionally, direct feedback from actual users of our system can serve as a
useful measure of its merits in practice.

Fortunately, virtual machine performance is not significantly affected by Data
Capsules, since it requires no significant changes to qemu or KVM.  Basic VM management functions also run quickly, with starting and stopping and deleting VMs taking only fractions of a second.  VM creation has also been made almost instantatneous to the user through clever use of copy-on-write references to base disk images.  This leaves only the performance of switches between modes.

For our evaluation, we plan to consider two useful measurements for assessing
the speed of a migration-based mode switch.  Since VM state begins to be
transferred to the target secure machine immediately upon startup, there is a
time from startup until the state transfer tapers to a steady stream as most of
the initial state is completely transferred and only new updates are then
carried over.  In our tests on two servers linked by a direct wired connection,
this switch \emph{startup time} took about 20 seconds for a base installation of
Ubuntu 10.04, which took about 3.5 GB of virtual disk space.  We note that the
entire virtual hard disk contents are not transferred, but only copy-on-write
changes to the virtual hard disk, as well as the entire contents of memory and
CPU registers.  This startup time is easily approximated by the actual execution
time of a normal, undelayed migration between two machines.

The second element is the actual \emph{execution time} of a switch, assuming
that the aforementioned startup process has already occurred and there is only a
small amount of update data that is being periodically generated and
transferred.  So long as the end user has used a machine for a time greater than
the startup time, and so long as bandwidth is not constrained so much that the
new data being generated cannot fully transfer within the window in which it is
generated, then there should only be a small, negligible delay, that is
virtually imperceptible to the user.

For our existing prototype, which does not use migration-based mode switches,
but rather makes use of VM live snapshots, it is not inaccurate to say that its
functionality is comparable to that of a migration-based mode switch without the
delay of transferring state across machines.  In this sense, the execution time
of a snapshot-based mode switch should certainly be no less than the execution
time of an equivalent migration-based execution time.  In light of this, we can
use the mode switch performance time for a snapshot-based mode switch to
approximate the upper bound for switch execution time in the final
migration-based Data Capsules system.  In practice, we have observed that this
switch time almost always falls within 5 seconds, and is usually closer to 1 or
2 seconds.

We also note for clarity that these switch times refer to the switch from
maintenance mode into secure mode.  Under a migration-based mode switch, the
switch back to maintenance mode simply requires unpausing a VM and switching
back the remote desktop or shell connection to the maintenance machine that VM
is running on.

@@

A user study is another important option for evaluating the undeniably
unfamiliar paradigm of having two machine modes for a single conceptual machine.
Although we have no plans to conduct a formal study in the near future, we are
in the process of implementing a prototype of the system that will be used by
humanities researchers for secure access to copyrighted materials.  After
initial alpha testing by a small number of subjects unfamiliar with the system,
feedback was positive, with no apparent complaints or confusion about how it was
used.  We leave more formal and in depth user studies for future work.

@@

In this paper, we introduced Data Capsules, a system that enables more usable
access to sensitive data repositories by eliminating the need to audit data and
software that are brought in to the secure environment.  Data Capsules presents
users with a maintenance mode for fully customizing their software environment,
and allows users to then import this environment into a secured environment
where they can then perform analysis on sensitive data with limited risk of data
leak.  Data Capsules achieves this mode paradigm by migrating underlying VMs
between machines segregated by their access to the network and to the sensitive
data.  A prototype of the system has been implemented, and initial performance
data and end user feedback are both promising.
